---
title: "Final_project"
author: "Josh VandenBerg"
date: "5/4/2020"
output: html_document
---
```{r}
d2018 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/2018_state.csv", skip = 12, header = FALSE, stringsAsFactors = FALSE)
d2016 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/2016_state.csv", skip = 12, header = FALSE, stringsAsFactors = FALSE)
d2014 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/2014_state.csv", skip = 12, header = FALSE, stringsAsFactors = FALSE)
d2013 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/2013_state.csv", skip = 14, header = FALSE, stringsAsFactors = FALSE)
d2012 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/2012_state.csv", skip = 15, header = FALSE, stringsAsFactors = FALSE)
d2010 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/10stall.csv", skip = 14, header = FALSE, stringsAsFactors = FALSE)
d2008 <- read.csv("/Users/joshuavandenberg/Documents/Data Manage/08stall.csv", skip = 13, header = FALSE, stringsAsFactors = FALSE)

names(d2018) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll")
names(d2016) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll")
names(d2014) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll")
names(d2013) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll")
names(d2012) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll")
names(d2010) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll", "Index")
names(d2008) <- c("State","Government_Function","FT_Employment","FT_Payroll", "PT_Employment","PT_Payroll","PT_Hours","FT_equiv_Employment","March_Employment","March_Payroll","Index")
d2018 <- subset(d2018, select = -c(FT_equiv_Employment, March_Employment, March_Payroll))
d2016 <- subset(d2016, select = -c(FT_equiv_Employment, March_Employment, March_Payroll))
d2014 <- subset(d2014, select = -c(FT_equiv_Employment,March_Employment, March_Payroll))
d2013 <- subset(d2013, select = -c(FT_equiv_Employment,March_Employment, March_Payroll))
d2012 <- subset(d2012, select = -c(FT_equiv_Employment,March_Employment, March_Payroll))
d2010 <- subset(d2010, select = -c(FT_equiv_Employment,March_Employment, March_Payroll, Index))
d2008 <- subset(d2008, select = -c(FT_equiv_Employment,March_Employment, March_Payroll,Index))
d2018$year <- 2018
d2013$year <- 2013
d2008$year <- 2008
d2010$year <- 2010
d2016$year <- 2016
d2012$year <- 2012
d2014$year <- 2014
full <- rbind(d2018,d2016,d2014,d2013,d2012,d2010,d2008)
full$Government_Function <- sub("Total - All Government Employment Functions","Total",full$Government_Function)
full$State <- sub("United States","US",full$State)
full$FT_Employment <- as.numeric(gsub(",","",full$FT_Employment))
full$PT_Employment <- as.numeric(gsub(",","",full$PT_Employment))
full$region <- with(full,ifelse(State %in% c("CA","AZ","AK","NM","HI","CO","WY","MT","ID","UT","NV","OR","WA","California","Arizona","Alaska","New Mexico","Hawaii","Colorado","Wyoming","Montana","Idaho","Utah","Nevada","Oregon","Washington"),"W",ifelse(State %in% c("TX","OK","AR","LA","MS","TN","AL","GA","FL","SC","NC","KY","WV","VA","MD","DE", "Texas","Oklahoma","Arkansas","Louisiana","Mississippi","Tennessee","Alabama","Georgia","Florida","South Carolina","North Carolina","Kentucky","West Virginia","Virginia","Maryland","Delaware"),"S",ifelse(State %in% c("IA","MN","ND","SD","NE","KS","MO","WI","IL","IN","MI","OH","Iowa","Minnesota","North Dakota","South Dakota","Nebraska","Kansas","Missouri","Wisconsin","Illinois","Indiana","Michigan","Ohio"),"MW",ifelse(State %in% c("ME","NH","VT","MA","RI","CT","NY","PA","NJ", "Maine","New Hampshire","Vermont","Massachusetts","Rhode Island","Connecticut","New York","Pennsylvania","New Jersey"), "NE","US")))))
total <- full[full$Government_Function == "Total",]
finance <- full[full$Government_Function == "Financial Administration",]
totalus <- total[total$State=="US",]
totalreg <- total[total$State!="US",]
totalmw <- total[total$region=="MW",]
totalS <- total[total$region=="S",]
totalW <- total[total$region=="W",]
totalne <- total[total$region=="NE",]
financeus <- finance[finance$State=="US",]
education <- full[full$Government_Function == "Education Total",]
educationus <- education[education$State=="US",]
hospital <- full[full$Government_Function == "Hospitals",]
hospitalus <- hospital[hospital$State=="US",]
judicial <- full[full$Government_Function == "Judicial and Legal",]
judicialus <- judicial[judicial$State=="US",]
police <- full[full$Government_Function == "Police Protection Total",]
policeus <- police[police$State=="US",]
highway <- full[full$Government_Function == "Highways",]
highwayus <- highway[highway$State=="US",]
mwtotal <- aggregate(FT_Employment~year,totalmw,mean)
stotal <- aggregate(FT_Employment~year,totalS,mean)
netotal <- aggregate(FT_Employment~year,totalne,mean)
wtotal <- aggregate(FT_Employment~year,totalW,mean)
mwtotalpt <- aggregate(PT_Employment~year,totalmw,mean)
stotalpt <- aggregate(PT_Employment~year,totalS,mean)
netotalpt <- aggregate(PT_Employment~year,totalne,mean)
wtotalpt <- aggregate(PT_Employment~year,totalW,mean)
```

```{r}
knitr::opts_chunk$set(echo = FALSE)
```
# Project
#### Joshua VandenBerg
#### May 6, 2020

## Intro

This project will be looking at the trend of government employees not only in different government functions but also in different regions of the Unites States. I will be looking at the years from 2008-2018. I did not chose every year instead I just used 2008, 2010, 2012, 2013, 2014, 2016, and 2018. Interstingly, the recession in 2012 is quite noticeable in all of the United States, but not in every government function. This project will also look into if there is a difference between full-time government employees compared to part-time.

One part of this project that did not get looked at is more in-depths state research. I focused more on regions of the United States, but it would be really interesting to see how different states reacted over time.

All of the data is ASPEP datasets from the United States Census Bureau. Each year was its separte file. So, I just binded all the data while making a year column, so that I could compare each year. I also had to make a region variable. I separated the states into four regions as the U.S. Census Bureau separated the states. The 2008 and 2010 datasets needed a little more cleaning as there was a few formatting changes compared to those of more recent years.

## Exploration

These first two graphs are just looking at the total number of government employees over the 10 year range. It is split between full-time employees and part-time employees.

```{r}
par(mfrow=c(1,2))
plot1 <- plot(totalus$year,totalus$FT_Employment, type = "l", col = "blue", main = "Number of Full-time Employees", xlab = "Year", ylab = "FT Employment")
plot2 <- plot(totalus$year,totalus$PT_Employment, type = "l", col = "red",main = "Number of Part-time Employees", xlab = "Year",ylab = "PT Employment")
```

The number of government full-time employees looks as if it was affected around the recession as there were more than 8,000 full time employees in 2008 to that of 2012. By 2018, it was close to the 2008 totals. On the other hand, the number of part-time employees has continued to increase. The next boxplots are looking at the different regions and the number of part-time or full-time employees.
```{r}
par(mfrow=c(1,2))
boxplot(totalreg$FT_Employment~totalreg$region, main = "Full-time Employees by Region", xlab = "Region", ylab = "FT Employment")
boxplot(totalreg$PT_Employment~totalreg$region, main = "Part-time Employees by Region", ylab = "PT Employment", xlab = "Region")
```

The first thing I noticed is the outliers in the South and West regions coming from Texas and California respectively. Of the full-time employees, the boxplots are very similar to populations of states in these regions. The MidWest is all very similar. The NorthEast is very spread out, especially with the likes of New York. The South is farely similar besides Texas. The West is actually not as populated compared to other states in the US besides California. The part-time employee boxplot on the other hand looks much similar. The box or middle 50% is around the same size to that of the other regions.

## Analysis
```{r}
par(mfrow=c(2,2))
plot(mwtotal, type = "l", col ="red", main = "Midwest", ylab = "",las=1)
plot(stotal, type = "l", col ="darkred", main = "South", ylab = "",las=1)
plot(netotal, type = "l", col ="darkblue", main = "Northeast", ylab = "",las=1)
plot(wtotal, type = "l", col ="blue", main = "West", ylab = "",las=1)
```

```{r}
par(mfrow=c(3,2))
plot(financeus$year,financeus$FT_Employment, type = "l", col = "red",ylab = "",las=1, xlab = "Year", main = "Finance")
plot(educationus$year,educationus$FT_Employment, type = "l", col = "orange",ylab = "",las=1,xlab = "Year", main = "Education")
plot(hospitalus$year,hospitalus$FT_Employment, type = "l", col = "darkred",ylab = "",las=1,xlab = "Year", main = "Hospital")
plot(judicialus$year,judicialus$FT_Employment, type = "l", col = "darkgreen",ylab = "",las=1,xlab = "Year", main = "Judicial")
plot(policeus$year,policeus$FT_Employment, type = "l", col = "darkblue",ylab = "",las=1,xlab = "Year", main="Police")
plot(highwayus$year,highwayus$FT_Employment, type = "l", col = "darkgrey",ylab = "",las=1,xlab = "Year",main="Highway")
```

```{r}
par(mfrow=c(2,2))
plot(mwtotalpt, type = "l", col ="red",main = "Midwest", ylab = "",las=1)
plot(stotalpt, type = "l", col ="darkred",main = "South", ylab = "",las=1)
plot(netotalpt, type = "l", col ="darkblue",main = "Northeast", ylab = "",las=1)
plot(wtotalpt, type = "l", col ="blue",main = "West", ylab = "",las=1)
```
```{r}
par(mfrow=c(3,2))
plot(financeus$year,financeus$PT_Employment, type = "l", col = "red",ylab = "",las=1, xlab = "Year", main = "Finance")
plot(educationus$year,educationus$PT_Employment, type = "l", col = "orange",ylab = "",las=1, xlab = "Year", main = "Education")
plot(hospitalus$year,hospitalus$PT_Employment, type = "l", col = "darkred",ylab = "",las=1, xlab = "Year", main = "Hospital")
plot(judicialus$year,judicialus$PT_Employment, type = "l", col = "darkgreen",ylab = "",las=1, xlab = "Year", main = "Judicial")
plot(policeus$year,policeus$PT_Employment, type = "l", col = "darkblue",ylab = "",las=1, xlab = "Year", main = "Police")
plot(highwayus$year,highwayus$PT_Employment, type = "l", col = "darkgrey",ylab = "",las=1, xlab = "Year", main = "Highway")
```
